---
title: "Week 6 Presentation"
author: "Steele H. Valenzuela"
date: "October 23, 2015"
output: slidy_presentation
---

## Let's Set Up Camp (do before class) [1 of 3]

Go to the following repo and download the following:
**github.com/shv38339/statcomp597d**

- *predictvals.R*
- *eqn.R*

## Let's Set Up Camp (do before class) [2 of 3]

Next, install and upload the following libraries:

```{r, messages = FALSE}
library(dplyr, warn.conflicts = FALSE)
library(ggplot2)
library(gcookbook)
library(nycflights13)
library(broom)
library(lubridate)
```

## Let's Set Up Camp (do before class) [3 of 3]

And rename the following data sets (*optional*)

```{r}
htwt <- heightweight # from the gcookbook library
flts <- flights # from the nycflights13 library
```

## Predicted Values [1 of 1]

What if you want to add your own prediction line to a plot? 

First, let's create a scatterplot using the `heightweight` data set (aptly renamed `htwt`). 
```{r}
# You may want to check out the data set first with the following command
sp <- ggplot(data = htwt, aes(x = ageYear, y = heightIn)) + 
      geom_point(colour = "grey40")
```

Now, let's create the corresponding linear model...
```{r}
model <- lm(heightIn ~ ageYear, data = htwt)
```

...and generate two new explanatory and response variables (aka X and Y).
```{r}
source(file = "predictvals.R") # from my github. save into your working directory
lin.pred <- predictvals(model, "ageYear", "heightIn")
```
When you reach the end, examine the first 6 values of `lin.pred`.
Additionally, examine the file *predictvals.R* to get a sense of what the function is doing. It's quite complex but understanding bits and pieces is crucial to see what's going on under the hood.

## Annotated Text to a plot [1 of 1]

How would you find the $R^2$ value of a linear model?
```{r, eval = FALSE}
str(summary(model))
```
So where was $R^2$? 

Now, let's explore the add-on function `annotate(...)` (*Hint: `?annotate` to read about it.)
```{r, eval = FALSE}
sp + annotate("text", label = "r^2 = ???", x = Inf, y = -Inf, hjust = 1.1, vjust = -.5)
```

Check out the file *eqn.R* and see how it creates the object **eqn**, which we will input into `annotate(...)`.
```{r, eval = FALSE}
source(file = "eqn.R")
sp + geom_line(data = lin.pred) +
      annotate("text", label = eqn, parse = TRUE, x = Inf, y = -Inf, hjust = 1.1, vjust = -.5)
```

## Exercise 1

1. Create a quadratic model from the `heightweight/htwt` data set.
 a. (Hint: `model <- lm(y ~ x + I(x^2), data = dat)`)
2. Create a scatterplot with a fitted quadratic model line.
3. Insert the coefficient values and $R^2$ value as annotated text (Hint: will have to edit file `eqn.R`).
4. Change the name of the axes and add a title.


## Quick Overview of Broom

How did we extract the $R^2$ value earlier?

Run the following functions on the object **model** that was created earlier.
- `tidy(...)`
- `augment(...)`
- `glance(...)`


## dplyr

Let's check out the flights data set
```{r, eval = FALSE}
dim(flights)
head(flights)
```
In base `R`, how would you find all flights on December 25th?


## dplyr Data Format

When working with `dplyr`, save the data set as follows:
```{r, eval = FALSE}
data.df <- tbl_df(data) 
```
There's also no harm in saving over the object itself...or is there?


## arrange(...) & select(...)


```{r, eval = FALSE}
arrange(flights, year, desc(month), desc(day))

select(flights, carrier:air_time)
select(flights, -(year:carrier))
```

## mutate(...)

```{r, eval = FALSE}
htwt1 <- mutate(htwt, bmi = 703*(weightLb/(heightIn^2)))
head(htwt1)

head(transmute(htwt, bmi = 703*(weightLb/heightIn^2)))
```


## summarise(...)

```{r, eval = FALSE}
summarise(flights, avg.dist = mean(distance))

gender <- group_by(htwt, sex)
summarise(gender, avg.Lb = mean(weightLb), avg.In = mean(heightIn))
```


## Pipe Operator %>%

Rather than writing nested code, we may now write code from left to right (or top to bottom).
```{r, eval = FALSE}
pipe1 <- htwt %>% 
          group_by(sex) %>% 
            summarise(avg.Lb = mean(weightLb), 
                      avg.In = mean(heightIn))
pipe1
```


## Nested functions

In `R` (and I'm sure in other languages), code is nested and can become quite messy and cumbersome to read, write, and execute. Here's an example.

```{r, eval = FALSE}
filter(
  summarise(
    select(
      group_by(flights, year, month, day),
      arr_delay, dep_delay
    ),
    arr = mean(arr_delay, na.rm = TRUE),
    dep = mean(dep_delay, na.rm = TRUE)
  ),
  arr > 30 | dep > 30
)
```


## Pipe Operator at full speed


```{r, eval = FALSE}
flights %>%
  group_by(year, month, day) %>%
    select(arr_delay, dep_delay) %>%
      summarise(arr = mean(arr_delay, na.rm = TRUE),
                dep = mean(dep_delay, na.rm = TRUE)) %>%
        filter(arr > 30 | dep > 30)
```


## For more info on dplyr

Here are more of the perks `dplyr` has to offer. This vignette will be extremely valuable. 
```{r, eval = FALSE}
browseVignettes(package = "dplyr")
```


## Homework (yay!)

See my repo on GitHub for week 6 homework!

