---
title: "Functions"
subtitle: "Are they (Un)necessary?"
author: "Steele H. Valenzuela - the best TA ever"
date: "November 16, 2015"
output: 
  ioslides_presentation:
  widescreen: true
---

# Quiz

## Basic Format of a Function {.smaller}

In `R`, our functions tend to be functions of the base functions.
```{r}
a <- sample(x = 10, size = 20, replace = TRUE)

hello <- function() cat("hello, world!\n") # necessary, of course

f1 <- function(x) x + 3 # one line functions

f2 <- function(x, y) { # more than one line functions
  x + y 
}                      # include these curly braces
```

What will be outputted (is that a verb?)?
```{r}
f1(10)
f2(10, -10)
```

## Why should I learn to write functions? (From Nick's slides)

* If repeating operations within your code, consider functionalizing it
* Pick good, short, clear, active names (particularly verbs)
* Creates standardized ways to perform certain operations
* Decreases likelihood of errors
* Less code is typed, especially when changing one piece
* Facilitates literate coding, making things simpler
* Makes debugging simpler

## Basic Format of a Function (cont.)

```{r function.name, eval = FALSE}
function.name <- function(arguments) {
  # body of a function
  ...
  return(something) # always include a return statement
}
```
Can you easily add functions in other statistical programs?


## You can communicate with the user

Often, these are couched in if-statements, i.e. "if some unusual condition is met, here is something you should know".
```{r error function, eval = TRUE}
error_msg_fun <- function(x) {
  message("Lift off!")
  warning("Houston, we have a minor glitch. No biggie.")
}
error_msg_fun(a)
```
Does it matter what we input into the function `error_msg_fun(...)`?


## You can communicate with the user (cont.) {.build}

Let's try `stop(...)`.
```{r error function 2, eval = TRUE, error=TRUE}
error_msg_stop <- function(x) {
  if(class(x) != "numeric")
    stop("x must be numeric!")
  a <- x^2 + x - 10
  return(a)
}
error_msg_stop("chicken")
error_msg_stop(3)
```

## Other functional features...

```{r my_fun, eval = TRUE, error = TRUE}
my_fun <- function(x, ...) {
  require(ggplot2)
  if(class(x) != "numeric" & class(x) != "integer")
    stop("x must be numeric")
  p <- qplot(x, ...)
  print(p)
}
```
Notice the `...`? What does it allow us to do? 

## Other functional features (cont.)... {.build .smaller}

First, let's check the class of *a*, from the first slide and then input *a* into our function.
```{r, message = TRUE, fig.height=3, fig.width=3, fig.align='center'}
my_fun(a)
```

Let's take advantage of the message we see and use the `...` feature. Change the *binwidth* option for this function.


## A flood functions that may be useful {.smaller}

* `class(...)` & `typeof(...)` for characters and numbers
* `names(...)` & `str(...)` to view the contents of an object
* `args(...)` to view the arguments of a function
* `length(...)` & `dim(...)` for the length of a vector or data frame/matrix
* `cat(...)`, `print(...)`, or `paste(...)` for text
* `!=`, `==`, `<=`, `=>`, `&`, `|` for logical operators
* `round(...)` for outputted numbers that you'd like to truncate
* `if` & `ifelse` for conditional statements
* `while` and `for` loops
* `as.numeric(...)`, `is.numeric(...)`, and others
* and many more of the base functions located at this [link][id1]

[id1]: https://cran.r-project.org/doc/contrib/Short-refcard.pdf


## Resources

Here are some of the resources I used for this lecture and others which cover more advanced areas of creating functions.

* Hadley's [page][id2]
* UCLA's [page][id3] (a bit on `for` loops and the `while` function)
* Prof. Nicholas Reich's [slides][id4]
* Official [R documentation][id5] (meticulous, but a necessary read)

[id2]: http://adv-r.had.co.nz/Exceptions-Debugging.html
[id3]: http://www.ats.ucla.edu/stat/r/library/intro_function.htm
[id4]: http://nickreich.github.io/statComp2014/lectures/week10/functions-in-R.pdf
[id5]: https://cran.r-project.org/doc/manuals/R-intro.html#Writing-your-own-functions


## Homework

1. Follow this [github workflow][id6] for the homework.
    * If you have previously "cloned" the **statcomp597d** repo onto your local computer, delete it and follow the github workflow. 
    * After you have "forked" the statcomp597d online repo...
    * Clone **your** online repo to your local machine...
    * Complete the homework...
    * Commit the changes...
    * Push the changes to your online repo...
    * And finally, create a pull request. 

[id6]: http://www.princeton.edu/~mjs3/soc504_s2015/submitting_homework.shtml





